---
title:    "Final Exam"
subtitle: "Bayesian Statistics"
author:   "Daniel Carpenter"
date:     "May 9, 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document:
    toc: yes
    toc_float: yes
    # number_sections: yes
    toc_depth: 2
  # github_document:
  #   toc: yes
  #   number_sections: yes
  #   toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo    = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE,
	cache   = TRUE
)
```


Please place all answers in this document - use $\LaTeX$, R and Jags as needed. Answers to many qestions will be made dynamically through an R chunk.

# Question 1 - 

## 1. (a)
## 1. (b)
## 1. (c)
## 1. (d)
## 1. (e)
## 1. (f)
### 1. f, i)
### 1. f, ii)

---

<br>

# Question 2

## 2. a)
## 2. b)
## 2. c)
## 2. d)
## 2. e)

---

<br>

# Question 3

## 3. a)
## 3. b) 
## 3. c)
## 3. d)
## 3. e)

---

<br>


# Question 4

## 4. a)
## 4. b) 
## 4. c)
## 4. d)
### 4. d, 1)
### 4. d, 2)

---

<br>


# Question 5 - Exponential Family

The exponential family can be defined as any density of the following form:
$$
f(y \mid \theta, \phi)=\exp \left(\frac{y \theta-b(\theta)}{a(\phi)}+c(y, \phi)\right)
$$

## 5. a) Show $E(Y) = b^{\prime}(\theta)$

<br>

## 5. b) Show $V(Y) = b^{\prime \prime}(\theta) a(\phi)$

<br>

## 5. c) Show that the `Binomial` distribution belongs to the `exponential` family

<br>

## 5. d) Show $E(Y) = np$ 

<br>

## 5. e) Show $V(Y) = np(1−p)$

---

<br>


# Question 6

---

<br>

# Question 7

## 7. a)
## 7. b)
## 7. c)
## 7. d)

---

<br>


# Question 8


## 8. a)
## 8. b)
## 8. c)
## 8. d)

---

<br>


# Question 9

```{r data9}
dataList = list(x = c(-10, -9, -8, -7, -6, -5, -4, -3, -2, -1, 0, 1, 2, 
3, 4, 5, 6, 7, 8, 9, 10), y = c(21.72, 20.41, 14.06, 14.6, 13.8, 
10.4, 8.22, 6.79, 4.92, 0.45, 4.57, 1.95, 10.14, 13.28, 21.88, 
18.88, 22.91, 23.53, 34.99, 35.07, 39.17), N = 21)
```

## 9. b)
### 9. b,i)
### 9. b,ii)
### 9. b,iii)

## 9. c)
## 9. d)
## 9. e)

---

<br>


# Question 10

```{r}
x=50:80
set.seed(24)
y=10+2*x+rnorm(31,0,10)

xx=-15:15
set.seed(24)
yy=10+2*xx+rnorm(31,0,10)

```

## 10. a)
## 10. b)
### 10. b, i)
### 10. b, ii)



