---
title: 'Lab 12: Titanic data'
subtitle: 'Bayesian Statitics'
author: "Daniel Carpenter"
date: "April 2022"
output:
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The course

As you have noticed "the Titanic" has been a theme of the course. The course has been structured into three parts:

    * Distributional results and basics of R
    * Binomial: a simple model where Bayesian methodology is learnt
    * The GLM: A more advanced application of Bayesian theory

All the skills that you have learnt in parts 1 and 2 will now be applied to a logistic regression which is a special case of the GLM.

We will now start to analyze the Titanic data set and you will perfect this in Assignment 4.

# Task 0: The story

## Summarize the Titanic story by reading the following web page: (https://www.history.com/this-day-in-history/unsinkable-titanic-sinks)

In 1912, the Titanic, the largest and fastest boat in the world that carried 2,200 passengers, sank in the Atlantic. 
The boat had 16 compartments; if the ship hit an iceberg, the incision must rupture 4 of the 16 
compartments to sink the ship. Propaganda followed, and many assumed the largest and fastest 
ship was unsinkable. One night, the ship soared across the Atlantic and met an iceberg head-on, 
breaching 5 or more of the compartments, which caused the boat to sink. The bow submerged first, then 
the boat broke in half and fully sank over 13,000 feet to the ocean floor. 700 of the 2,200 passengers died due to insufficient 
lifeboats and other precautionary measures. Lifeboats prioritized women and children, who made up most 
of the surviving members.


# Task 1: The Titanic data set
We will use the data set as prepared in the `vcdExtra` package. We will aslo use the `gpairs` package.
Please install.

##  Make a pairs plot of the data using `gpairs()`
```{r data, message = FALSE, warning=FALSE}
require(vcdExtra) # Titanic passenger data
require(gpairs)   # Use the data frame below and use the function gpairs()

data("Titanicp")
head(Titanicp)

# Huge summary of the dataset
gpairs(Titanicp)
```

Notice that there are a number of categorical variables and a continuous variable `age`.

## Using the R help for the package describe the variables in the Titanicp data set.

# Task 2: Interpreting the plots

## Interpret the plots below:

```{r ggplots}
library(ggplot2)

# 
g = ggplot(Titanicp, aes(x = age, 
                         y=as.numeric(survived =="survived"),
                         color = sex)) + 
  ylab("Survived") + 
  theme_bw() 

# Jitter the plot so that the points are not on top of each other
# Mainly woman who survived, but not exact relationship
JITTER_AMOUNT = 0.015
g = g + geom_point(position = position_jitter(height = JITTER_AMOUNT, width =0))
g

# Generalized Linear Model - binomial with link logit. Also fit a cubic
# Older woman are more likely to survive, and younger boys are more likely to survive
# Note wide interval likely due to limited sample size
g = g + stat_smooth(method = "glm", 
                    method.args = list(family=binomial("logit")), 
                    formula = y ~ x + I(x^2) + I(x^3), 
                    alpha = 0.25, size = 2.5, 
                    aes(fill = sex))
g

# Now break out by class of passenger
# First class females have high likelihood of survival
g = g + facet_wrap(~pclass)
g


```




# Task 3: Classical analysis using `glm()`

We will perform a logistic regression using glm.  

* Need maximum liklihood estimates (`Estimate`) for the MCMC using glm
* Interaction used because the survival depends what level of age you are at or if you are male or female
* Note very significant
* Goes through 4 chains

```{r classicalglm, eval=TRUE}
clglm = glm(survived ~ sex + age + sex:age,family = "binomial", data = Titanicp)
summary(clglm)

```

## What are the classical point estimates?


# Task 4: Use the classical model to make data for Jags

## Complete the code below (one line, `y=`)


## Why not just use the original data?:
* Get rid of the `NAs`
* Also the model matrix is easy to handle
* The original data uses boolean text to describe if the passenger survived, so need to convert to binary ($1$, $0$) to model in R

```{r datamodel, eval=TRUE}
mat1=model.matrix(clglm)
mat2=model.frame(clglm)
head(mat1)
head(mat2)

# Create output
y = with(mat2, ifelse(survived == "survived", 1, 0))

# Data list for MCMC model
dataList=list(y = y, x = mat1[, "age"],sexm = mat1[,"sexmale"], sexmx = mat1[,"sexmale:age"] , n = length(y))

```



# Task 5: Now we will use the classical estimates as initial values in a jags script

## Complete the Jags script below

### Warning
The MCMC sampler is very sensitive to initial values. Within each gibbs iteration Jags will choose a slice sampler -- this will take some time. You may need to wait a few minutes for the sampling to complete.

```{r jagscode, warning=FALSE,message=FALSE,eval = TRUE}

library(rjags)

#Define the model:
modelString = "
model{
    for(i in 1:n) {
      y[i] ~ dbin(theta[i], 1)
      logit(theta[i]) <- beta[1] + beta[2]*x[i] + beta[3]*sexm[i] + beta[4]*sexmx[i]
    }
    
  for(j in 1:4){
    beta[j] ~ dnorm(0,1.0E-3)
  }
}
"
writeLines( modelString , con="TEMPmodel.txt" )

initsList = list(beta = c(0.5,0.02,-1.15,-0.05)) # initial values

# Run the chains:
jagsModel = jags.model( file="TEMPmodel.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )
list.samplers(jagsModel)

update( jagsModel , n.iter=500 )
codaSamples = coda.samples( jagsModel , variable.names=c("beta"),
                            n.iter=33340 )
save( codaSamples , file=paste0("lab12","Mcmc.Rdata") )



library(ggmcmc)
s = ggs(codaSamples)
d=ggs_density(s)

print(d)

cr =  ggs_crosscorrelation(s)
print(cr)

summary(codaSamples)

```


# Task 6: Interpretation of MCMC

## Interpret all the Bayesian output 
* All values stated as approximately since each MCMC run will generate a slightly different estimate.

### Interpret the point estimates for the betas
$\beta_1$: the mean value ($\mu$) for $\beta_1$ is around  0.50952 and the $\sigma$ is 0.257209. I.e., each person could either survive   
$\beta_2$: the mean value ($\mu$) for $\beta_2$ is around  0.02229 and the $\sigma$ is 0.008631 I.e. the older you are, the less likely you are to survive.   
$\beta_3$: the mean value ($\mu$) for $\beta_3$ is around -1.17196 and the $\sigma$ is 0.344398 I.e. a male is more likely to die than a female.   
$\beta_4$: the mean value ($\mu$) for $\beta_4$ is around -0.04620 and the $\sigma$ is 0.011349 I.e. if you are a male, your outcomes worsen even more if as you age    

### Interpret the interval estimates for the betas
* All values stated as approximately since each MCMC run will generate a slightly different estimate.

With a probability of 95%, the mean value for:
$\beta_1$ will be between  0.006657 and 1.01637  
$\beta_2$ will be between  0.005433 and 0.03958  
$\beta_3$ will be between -1.860702 and -0.49767  
$\beta_4$ will be between -0.068774 and -0.02388  

### How do you know the MCMC  sampler converged to stationarity?
It is likely that the sampler converged since each chain well represents each other.

### Compare your results with the classical analysis estimates
For each variable, the MCMC estimate represents the classical estimate nearly exactly
