---
title:    "Assignment 2"
subtitle: "Bayesian Statistics"
author:   "Daniel Carpenter"
date:     "March 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

---

# Task `1`

## `a` Bayes Thereom
$$
\underbrace{p(\theta \mid D)}_{\text {posterior }}
=\underbrace{p(D \mid \theta)}_{\text {likelihood}} \times
\underbrace{p(\theta)}_{\text {prior}} 
/ \underbrace{p(D)}_{\text {evidence }}
$$

## `b` General Posterior
> Derive the general posterior result for a Beta prior and Binomial likelihood  

*Taken from JK's book - page 132 Doing Bayesian Data Analysis:* <br>

#### Bayes' rule
$p(\theta \mid z, N)=p(z, N \mid \theta) p(\theta) / p(z, N)$ <br>

#### Define Bernoulli and beta distributions
$=\theta^{z}(1-\theta)^{(N-z)} \frac{\theta^{(a-1)}(1-\theta)^{(b-1)}}{B(a, b)} / p(z, N)$ <br>

#### Rearrange factors
$=\theta^{z}(1-\theta)^{(N-z)} \theta^{(a-1)}(1-\theta)^{(b-1)} /[B(a, b) p(z, N)] \quad$ <br>

#### Collect Powers
$=\theta^{((z+a)-1)}(1-\theta)^{((N-z+b)-1)} /[B(a, b) p(z, N)] \quad$ <br>
$=\theta^{((z+a)-1)}(1-\theta)^{((N-z+b)-1)} / B(z+a, N-z+b)$ <br>


## `c` Plot Prior, Likelihood, and Posterior
* Given uniform prior (for a Beta prior and Binomial likelihood)
```{r 1c, message=FALSE, warning=FALSE}
source("DBDA2E-utilities.R")  # Load definitions of graphics functions etc.
source("BernBeta.R")          # Load the definition of the BernBeta function

# Specify the prior:
t = 4              # Specify the prior MODE.
n = 10               # Specify the effective prior sample size.
a = 1                # Convert to beta shape parameter a.
b = 1                # Convert to beta shape parameter b.

Prior = c(a,b)       # Specify Prior as vector with the two shape parameters.

# Specify the data:
N = 10                         # The total number of flips.
z = 4                         # The number of heads.
Data = c(rep(0,N-z),rep(1,z))  # Convert N and z into vector of 0's and 1's.


# Plot the three graphs
posterior = BernBeta( priorBetaAB=Prior, Data=Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

## `d` `Mybeta()` 
* Function to Plot Prior, Likelihood, and Posterior
* Beta prior and Binomial likelihood  
```{r 1d, message=FALSE, warning=FALSE}

# Create flexible function that plots the Prior, Likelihood, and Posterior
# Given a Beta prior and Binomial likelihood   
Mybeta <- function(alpha = 1, beta = 1, n, x) {

  source("DBDA2E-utilities.R")  # Load definitions of graphics functions etc.
  source("BernBeta.R")          # Load the definition of the BernBeta function
  
  # Specify the prior:
  t = x                          # Specify the prior MODE.
  
  Prior = c(alpha,beta)          # Specify Prior as vector with the two shape parameters.
  
  # Specify the data:
  N = n                          # The total number of flips.
  z = x                          # The number of heads.
  Data = c(rep(0,N-z),rep(1,z))  # Convert N and z into vector of 0's and 1's.
  
  # Make plots for prior, lik, and post  
  posterior = BernBeta( priorBetaAB=Prior, Data=Data , plotType="Bars" , 
                        showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
}
```


## `e` Call `Mybeta()` given various paramters
```{r 1e, message=FALSE, warning=FALSE}
  Mybeta(alpha=2, beta=2, n=10, x=6)
  Mybeta(alpha=4, beta=2, n=10, x=6)
  Mybeta(alpha=2, beta=4, n=10, x=6)
  Mybeta(alpha=20, beta=20, n=10, x=6)
```

## `f` `Mybeta2()`
* Adjusted function `Mybeta2()` that produces the same plots as `mybeta()` 
* but releases command line Bayesian point estimates and BCI’s of whatever equal tail size
we wish – all estimating `p`
```{r 1f, message=FALSE, warning=FALSE}
Mybeta2 <- function(alpha = 1, beta = 1, n, x, alphalevel=0.05) {

  source("DBDA2E-utilities.R")  # Load definitions of graphics functions etc.
  source("BernBeta.R")          # Load the definition of the BernBeta function
  
  # Specify the prior:
  t = x                          # Specify the prior MODE.
  
  Prior = c(alpha,beta)          # Specify Prior as vector with the two shape parameters.
  
  # Specify the data:
  N = n                          # The total number of flips.
  z = x                          # The number of heads.
  Data = c(rep(0,N-z),rep(1,z))  # Convert N and z into vector of 0's and 1's.
  
  # Make plots for prior, lik, and post  
  posterior = BernBeta( priorBetaAB=Prior, Data=Data , plotType="Bars" , 
                        showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
  
  # Get alpha and beta used to form the posterior distribution
  postAlpha = posterior[1]
  postBeta  = posterior[2]
  
  # Calculate the HDI Tails
  lowerLevel = (alphalevel / 2)
  upperLevel = (1 - alphalevel / 2)
  HDI = qbeta(c(lowerLevel, upperLevel), postAlpha, postBeta)
  
  # Return the Lower and Upper HDI at the given level of alpha
  return(list('lowerHDI' = HDI[1],
              'upperHDI' = HDI[2]))
}
```


* *Note that that R Markdown will not recreate the graphs since they are the same, but the `CMD` output is shown*
```{r 1fCall, message=FALSE, warning=FALSE}
# Call the above function under various parameter scenarios
Mybeta2(alpha = 2, beta=2, n=10, x=4, alphalevel=0.05) 
Mybeta2(alpha = 2, beta=2, n=10, x=4, alphalevel=0.10) 
Mybeta2(alpha = 2, beta=2, n=10, x=3, alphalevel=0.05)
Mybeta2(alpha = 2, beta=2, n=10, x=3, alphalevel=0.01)
```


# Task `2`

## `a` Show that mixbeta satisfies the first 2/3 properties of a density. 

### Assume the following:

#### 1. Two beta distributions with two shape parameters
$$
dbeta1 = dbeta(x, a1, b1) \\
dbeta2 = dbeta(x, a2, b2)
$$

#### 2. The sum of w is 1
$$
\sum_{-\infty}^{+\infty} w = 1
$$

### Explanation
i. $f(x) \geq 0$
  * Answer: The area under a beta distribution is from 0 to 1. Since the sum of the weight $w$ is 1, the area of both weighted betas will be 1.
ii. $\int_{-\infty}^{+\infty} f(x) d x=1$
  * Answer: If the sum of the weight $w$ is 1, and the area under a beta distribution must always be 1, then the area from -$\infty$ to +$\infty$ will cover that interval and sum to 1, making $\int_{-\infty}^{+\infty} f(x) d x=1$


## `b` `mixbeta()`
> Make an R function that will create the mixture density using `mixbeta()`

```{r 2b, message=FALSE, warning=FALSE}
mymix<- function(x, w, a1,b1, a2,b2) {
    w*dbeta(x, a1, b1) +(1-w)*dbeta(x, a2,b2)
  }
```

## `c` `Mymixplot()`
> Make a function that will plot the mixture

```{r 2c, message=FALSE, warning=FALSE}
Mymixplot <- function(w, a1,b1, a2,b2) {
  
  # Using mymix() to estimate mixed beta
  mymix<- function(x, w, a1,b1, a2,b2) {
    w*dbeta(x, a1, b1) +(1-w)*dbeta(x, a2,b2)
  }
  
  # Plot the curve and area with a title
  polygon(curve(mymix(x, w, a1,b1, a2,b2)), col = "skyblue3")
  title('Mixed Beta Plot\nDaniel Carpenter')
}

# Call the above function over various scenarios
Mymixplot(w=0.3, a1=2,b1=4, a2=4,b2=2) 
Mymixplot(w=0.5, a1=2,b1=4, a2=4,b2=2) 
Mymixplot(w=0.7, a1=2,b1=4, a2=4,b2=2)
Mymixplot(w=0.9, a1=2,b1=4, a2=4,b2=2)

```

## `d`

### `i` Analytical Posterior
$$
\underbrace{\frac{z+a}{N+a+b}}_{\text{posterior}} = \underbrace{\frac{z}{N}}_{\text{data}} \underbrace{\frac{N}{N+a+b}}_{\text{weight}} + \underbrace{\frac{a}{a+b}}_{\text{prior}} \underbrace{\frac{a+b}{N+a+b}}_{\text{weight}} .
$$

### `ii` Posterior Mixing Weight
* Posterior Mixing Weight is the weights applied to the two beta distributions to mix them

### `iii` Use a mixed beta as the prior  

*Note that the code below is adapted from JK's `BernBeta()` function - adjusted to take mixed beta priors*

```{r 2d, message=FALSE, warning=FALSE}
Mymixbeta <- function(w=0.5, n=10, x, a1=1, a2=1, b1=1, b2=1) {
  # Specify the prior:
  t = x                          # Specify the prior MODE.
  
  # Specify the data:
  N = n                          # The total number of flips.
  z = x                          # The number of heads.
  Data = c(rep(0,N-z),rep(1,z))  # Convert N and z into vector of 0's and 1's.
  
  # Create summary values of Data:
  z = sum( Data ) # number of 1's in Data
  N = length( Data ) 
  
  Theta = seq(0.001,0.999,by=0.001)                                                 # points for plotting
  pTheta = w*dbeta(Theta,a1,b1) +(1-w)*dbeta(Theta,a2,b2)                           # prior for plotting
  pThetaGivenData = w*dbeta(Theta, a1+z, b1+N-z) + (1-w)*dbeta(Theta, a2+z, b2+N-z) # posterior for plotting
  pDataGivenTheta = Theta^z * (1-Theta)^(N-z)                                       # likelihood for plotting
  
  
  # Plot Layout
  layout( matrix( c( 1,2,3 ) ,nrow=3 ,ncol=1 ,byrow=FALSE ) ) # 3x1 panels
  par( mar=c(3,3,1,0) , mgp=c(2,0.7,0) , mai=c(0.5,0.5,0.3,0.1) ) # margins
  cexAxis = 1.33
  cexLab = 1.75
  # convert plotType to notation used by plot:
  plotType="h"
  dotsize = 5 # how big to make the plotted dots
  barsize = 5 # how wide to make the bar lines    
  # y limits for prior and posterior:
  yLim = c(0,1.1*max(c(pTheta,pThetaGivenData)))
  
  
  # Plot the Prior
  plot( Theta , pTheta , type=plotType , 
        pch="." , cex=dotsize , lwd=barsize ,
        xlim=c(0,1) , ylim=yLim , cex.axis=cexAxis ,
        xlab=bquote(theta) , ylab='Mixed Beta' , 
        cex.lab=cexLab ,
        main="Prior (beta) - Daniel Carpenter" , cex.main=1.5 , col="skyblue")
  
  # Plot the likelihood
  plot( Theta , pDataGivenTheta , type=plotType , 
        pch="." , cex=dotsize , lwd=barsize ,
        xlim=c(0,1) , ylim=c(0,1.1*max(pDataGivenTheta)) , cex.axis=cexAxis ,
        xlab=bquote(theta) , ylab=bquote( "p(D|" * theta * ")" ) , 
        cex.lab=cexLab ,
        main="Likelihood (Bernoulli)" , cex.main=1.5 , col="skyblue" )
  
  # Plot the posterior.
  plot( Theta , pThetaGivenData , type=plotType , 
        pch="." , cex=dotsize , lwd=barsize ,
        xlim=c(0,1) , ylim=yLim , cex.axis=cexAxis ,
        xlab=bquote(theta) , ylab='Mixed Beta Posterior' , 
        cex.lab=cexLab ,
        main="Posterior (beta)" , cex.main=1.5 , col="skyblue" )
}

# Call the function and print the prior (mixed beta), likelihood and posterior)
Mymixbeta(w=0.5, n=10, x=4, a1=2, a2=2, b1=4, b2=4)
```
