---
title:  "Lab 2 - Bayesian Statistics - Distributional Theory"
author: "Daniel Carpenter"
date:   "January 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
  github_document: default
  pdf_document: 
    toc: TRUE
    number_sections: TRUE
  word_document:
    toc: TRUE
    number_sections: TRUE    
  md_document: default
     
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Binomial Distribution

### `LaTex` Formula for Discrete Binomial Distribution
> *Note* that the discrete binomial distribution is very useful for modeling processes in which the binary outcome can be either a success (1) or a failure (0)

$$
p(X = x|n, p) = 
  \begin{pmatrix}
  N \\
  k 
  \end{pmatrix}
p^k(1-p)^{N - k}
$$

### Function (`dmybin`) to Calculate Discrete Binomial Distribition
```{r dmybin, echo=TRUE}
dmybin <- function(X, n, p) {
  
  # Change X to k to be consistent with textbook
  k = X
  
  # Calculate the binomial coefficient (N k)
  binomialCoefficient <- choose(n, X)
  
  # Return the discrete binomial calculation
  return(binomialCoefficient * p^k * (1 - p)^(n - k))
}
```

### Call and Return Results of the `dmybin` Function
```{r call_dmybin, echo=TRUE}
y.dmybin = dmybin(X=0:4, n=10, p=0.5)
y.dmybin
```

### Call and Return Results of Base R Binomial Distribution Function `y.dbinom`
```{r call_dbinom, echo=TRUE}
y.dbinom = dbinom(x=0:4, size=10, prob=0.5)
y.dbinom
```

---

### Create a Cumulative Probability Function Called `pmybin`
```{r pmybin, echo=TRUE}
pmybin <- function(dmybin, x, n, p) {
  
  # Return the Cumulative Probability
  return(sum(dmybin(0:x, n, p)))
}
```

### Call and Return Results of the `pmybin` 
```{r call_pmybin, echo=TRUE}
cumulativeProbability.pmybin <- pmybin(dmybin, x=5, n=10, p=0.5)
cumulativeProbability.pmybin
```

### Call and Return Results of Base R Binomial Function Distribution Function `pbinom`
```{r pbinom, echo=TRUE}
cumulativeProbability.pbinom <- pbinom(q = 5, size=10, prob=0.5)
cumulativeProbability.pbinom
```

---

# Four Basic Distributional Functions: `dstem`, `pstem`, `rstem`, and `qstem`

```{r dists, echo=TRUE}

```


