---
title:  "Lab 2 - Bayesian Statistics - Distributional Theory"
author: "Daniel Carpenter"
date:   "January 2022"
fontsize: 12pt
geometry: margin=1in
output:
  md_document: default
     
  github_document: default
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
  pdf_document: 
    toc: TRUE
    number_sections: TRUE
  word_document:
    toc: TRUE
    number_sections: TRUE    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Discrete Binomial Distribution
> *Note* that the discrete binomial distribution is very useful for modeling processes in which the binary outcome can be ***either*** a success (`1`, `TRUE`) or a failure (`0`, `FALSE`)

### `LaTex` Formula for Discrete Binomial Distribution

$$
p(X = x|n, p) = 
  \begin{pmatrix}
  N \\
  k 
  \end{pmatrix}
p^k(1-p)^{N - k}
$$

### Function (`dmybin`) to Calculate Discrete Binomial Distribition
```{r dmybin, echo=TRUE}
dmybin <- function(X, n, p) {
  
  # Change X to k to be consistent with textbook
  k = X
  
  # Calculate the binomial coefficient (N k)
  binomialCoefficient <- choose(n, X)
  
  # Return the discrete binomial calculation
  return(binomialCoefficient * p^k * (1 - p)^(n - k))
}
```

### Call and Return Results of the `dmybin` Function
```{r call_dmybin, echo=TRUE}
y.dmybin = dmybin(X=0:4, n=10, p=0.5)
y.dmybin
```

### Call and Return Results of `base` R Binomial Distribution Function `dbinom`
```{r call_dbinom, echo=TRUE}
y.dbinom = dbinom(x=0:4, size=10, prob=0.5)
y.dbinom
```

---

### Create a Cumulative Probability Function Called `pmybin`
```{r pmybin, echo=TRUE}
pmybin <- function(dmybin, x, n, p) {
  
  # Return the Cumulative Probability
  return(sum(dmybin(0:x, n, p)))
}
```

### Call and Return Results of the `pmybin` 
```{r call_pmybin, echo=TRUE}
cumulativeProbability.pmybin <- pmybin(dmybin, x=5, n=10, p=0.5)
cumulativeProbability.pmybin
```

### Call and Return Results of `base` R Binomial Function Distribution Function `pbinom`
```{r pbinom, echo=TRUE}
cumulativeProbability.pbinom <- pbinom(q = 5, size=10, prob=0.5)
cumulativeProbability.pbinom
```

### Create a Binomial Distribution Plot
```{r bdist_plot, echo=TRUE}
x = 0:10

plot(x, 
     y = dbinom(x, size = 10, prob = 0.5),
     type = 'h', # h := histogram like
     main = 'Daniel Carpenter - Binomial Distributon',
     xlab = 'x')
```

---

# Poisson and Four Basic Distributional Functions: `dpois`, `ppois`, `rpois`, and `qpois`

## `a` Poisson Calculations

### Find ğ‘ƒ(ğ‘‹ = 4|ğœ† = 3)
* What is the probability that there are exactly 4 successes when 3 is the average?
```{r i, echo=TRUE}
dpois(x = 4, lambda = 3)
```
### Find ğ‘ƒ(ğ‘‹ â‰¤ 4|ğœ† = 3)
* What is the probability that there are 4 or less successes when 3 is the average? 
```{r ii, echo=TRUE}
ppois(q = 4, lambda = 3)
```
        
### Find ğ‘ƒ(ğ‘‹ > 4|ğœ† = 3)
* What is the probability that there are more than 4 successes when 3 is the average?
```{r iii, echo=TRUE}
ppois(q = 4, lambda = 3, lower.tail = FALSE)
```
      
### Find x so that ğ‘ƒ(ğ‘‹ â‰¤ ğ‘¥|ğœ† = 3)= 0.9997077 
* How many successes when 3 on average and cumulative probability of 0.9997077?
```{r iv, echo=TRUE}
qpois(p = 0.9997077, lambda = 3)
```
        
### Create a sample of size 100 from a Poisson distribution that has parameter ğœ† = 3. Store in an object.
```{r v, echo=TRUE}
poissonSample3 <- rpois(n = 100, lambda = 3)
```
        
### Make a second sample of size 100 from a Poisson that has parameter ğœ† = 6, store in an object
```{r vi, echo=TRUE}
poissonSample6 <- rpois(n = 100, lambda = 6)
```

## `b/c` Data Frame and Base `ggplot` for Boxplots and Violins
```{r b/c, echo=TRUE, message=FALSE}
if(!require(tidyverse)) install.packages(tidyverse)

# Create data frame with both samples
df <- data.frame(Fst = poissonSample3,
                 Snd = poissonSample6) %>%
  
  # Pivot data into single column for ggplot use
  pivot_longer(cols      = c("Fst", "Snd"),
               names_to  = "Sample",
               values_to = "x")


# Create a base Plot Object for future distribution graphs
basePlot <- ggplot(df,
                   aes(x = Sample,
                       y = x,
                       fill = Sample)) + 
            
            # Color palette and theme
            scale_fill_brewer(palette = "Pastel1") +
            theme_minimal() +
          
            # Title
            labs(title    = 'Sample Distribution Created from rpois Poisson Function',
                 subtitle = 'Sample Size (n) = 100 | Fst Lambda = 3; Snd Lamda = 6')
```
        
                
## `b` Create Box Plots
```{r boxplot, echo=TRUE}
basePlot + geom_boxplot()
```
                    
## `c` Create Violin Plots
```{r violin, echo=TRUE}
basePlot + geom_violin()
```
