---
title: "Lab 8 - Dummy variables and ANCOVA"
subtitle: "Bayesian Statistics"
author:   "Daniel Carpenter"
date:     "March 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


> Use an interaction model to predict salary based on wage and gender  
> Goal could be to identify discrepencies in linear models, which magnify changes over time.

The data was taken from:

[https://onlinecourses.science.psu.edu/stat502/node/188](https://onlinecourses.science.psu.edu/stat502/node/188)


[https://onlinecourses.science.psu.edu/stat502/node/187](https://onlinecourses.science.psu.edu/stat502/node/187)

```{r}
df = read.csv("salary-unequal.csv")
head(df)

df2 = read.csv("salary-equal.csv")
head(df2)
```


We can analyze the unequal slopes by adjusting and then using the following jags script

<br> 

---

# Task 1: (Unequal)

  Remember to take the `eval=FALSE` off the chunk once you are ready. Please fill in the following Jags script and then run it.

```{r jags,warning=FALSE, message=FALSE, eval=TRUE}
require(rjags)               # Must have previously installed package rjags.
fileNameRoot="tut12" # For output file names.

#df = read.table(file="salary-unequal.txt", sep = "\t", header =TRUE)
df = read.csv("salary-unequal.csv")
df
salary = df$salary
gender = df$gender
years = df$years
GM = ifelse(gender == "Male", 1,0)




Ntotal = length(salary)  # Compute the total number of data rows
dataList = list(    # Put the information into a list.
  years = years,
  salary = salary ,
  GM = GM,
  Ntotal = Ntotal 
)

#Define the model:
modelString = "
model{
  for(i in 1:Ntotal)
  {
    mu[i]<- beta0 + beta1*years[i] + beta2*GM[i] + beta3*years[i]*GM[i]
    salary[i] ~ dnorm(mu[i], tau)
  }
  
  beta0 ~ dnorm(0.0, 1.0E-6)
  beta1 ~ dnorm(0.0, 1.0E-6)
  beta2 ~ dnorm(0.0, 1.0E-6)
  beta3 ~ dnorm(0.0, 1.0E-6)
  sigma ~ dunif(0, 1000)
  tau <- pow(sigma,  -2)
}

" # close quote for modelString
writeLines( modelString , con="TEMPmodel.txt" )


#  initsList = list( theta=thetaInit )

initsList = list(beta0 = 0, beta1 = 0, beta2=0, beta3=0, sigma =10)

# Run the chains:
jagsModel = jags.model( file="TEMPmodel.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )
update( jagsModel , n.iter=500 )
codaSamples = coda.samples( jagsModel , variable.names=c("beta0", "beta1", "beta2" ,"beta3","sigma") ,
                            n.iter=33340 )
save( codaSamples , file=paste0(fileNameRoot,"Mcmc.Rdata") )

summary(codaSamples)

library(ggmcmc)
s = ggs(codaSamples)
ggs_density(s)

ggs_crosscorrelation(s)

```

# Task 2: (Unequal)

  Find parameter point and interval estimates. Interpret them!
```{r 2, eval=TRUE}
est.lm <- lm(salary ~ years + GM + years:GM)
summary(est.lm)
```
  
  
## Use these questions to help you in making interpretations in this lab:

#### 1. If `x` is increased by one unit what happens to the  mean value of `y`?
a. For each year increased, a male is paid the value of $beta3$ (which on the iteration before this was 10.098, it changes since it is MCMC) more than a female (thousand dollars? units not known since link is broken to data).
  b. E.g., in year 2, a male will make 10.098 $\times$ 2 more than the female on average. 
  
#### 2. Will the lines ever intersect over the range of the data?
a. No, the lines will never intersect
  
#### 3. What meaning does the slope have?
a. The slope of the interaction term means the amount more that is paid to males than females, given years of experience. 
b. The slope of years is the amount paid to males, given years of experience.
c. This should hover around 10. See $beta3$ normal density plot to get a sense of the potential variation.
  
#### 4. You can answer Q.1 using point and interval estimates.
a. For each year increased, a male is paid 10.00 more than a female.
  
#### 5. Make sure you know how to interpret BCI's (Bayesian credible intervals) -- these are probability intervals and NOT Confidence Intervals.
a. There is a 95% probability that a male is paid 14.210 or less than a female (since MCMC likely to vary from above output).

# Task 3: (Unequal)

  Plot the data and the estimating lines
```{r 3, eval=TRUE}
# Plot salary on years of experience
plot(salary~years)

# List of coefficients from model above (classical)
cf2 = coef(est.lm)

# Plot the slope of each male and female
abline(coef = c(cf2[1],cf2[2]), col="darkseagreen4", lwd =3)
abline(coef = c(cf2[1]+cf2[3], cf2[2]+cf2[4]), col = "steelblue", lwd = 3)
legend("topleft",legend = c("Female","Male"),fill =c("darkseagreen4","steelblue"))
title("Male vs. Female Salaries - Unequal Dataset | Daniel Carpenter")
```
  


<br>

---

# Task 4: (Equal)

Repeat task 1 - 3 for the ***equal*** slopes data and interpret the output. How should you change the model?

## Task 1: (Equal)

  Remember to take the `eval=FALSE` off the chunk once you are ready. Please fill in the following Jags script and then run it.

```{r jags2,warning=FALSE, message=FALSE, eval=TRUE}
fileNameRoot="tut12" # For output file names.

#df.equal = read.table(file="salary-unequal.txt", sep = "\t", header =TRUE)
df.equal = read.csv("salary-equal.csv")
df.equal
salary.equal = df.equal$salary
gender.equal = df.equal$gender
years.equal = df.equal$years
GM.equal = ifelse(gender.equal == "Male", 1,0)




Ntotal.equal = length(salary.equal)  # Compute the total number of data rows
dataList = list(    # Put the information into a list.
  years = years.equal,
  salary = salary.equal ,
  GM = GM.equal,
  Ntotal = Ntotal.equal 
)

#Define the model:
modelString = "
model{
  for(i in 1:Ntotal)
  {
    mu[i]<- beta0 + beta1*years[i] + beta2*GM[i] + beta3*years[i]*GM[i]
    salary[i] ~ dnorm(mu[i], tau)
  }
  
  beta0 ~ dnorm(0.0, 1.0E-6)
  beta1 ~ dnorm(0.0, 1.0E-6)
  beta2 ~ dnorm(0.0, 1.0E-6)
  beta3 ~ dnorm(0.0, 1.0E-6)
  sigma ~ dunif(0, 1000)
  tau <- pow(sigma,  -2)
}

" # close quote for modelString
writeLines( modelString , con="TEMPmodel.equal.txt" )


#  initsList = list( theta=thetaInit )

initsList = list(beta0 = 0, beta1 = 0, beta2=0, beta3=0, sigma =10)

# Run the chains:
jagsModel.equal = jags.model( file="TEMPmodel.equal.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )
update( jagsModel.equal , n.iter=500 )
codaSamples.equal = coda.samples( jagsModel.equal , variable.names=c("beta0", "beta1", "beta2" ,"beta3","sigma") ,
                            n.iter=33340 )
save( codaSamples.equal , file=paste0(fileNameRoot,"Mcmc.Rdata") )

summary(codaSamples.equal)

library(ggmcmc)
s = ggs(codaSamples.equal)
ggs_density(s)

ggs_crosscorrelation(s)

```

## Task 2 (Equal)

  Find parameter point and interval estimates. Interpret them!
```{r 2b, eval=TRUE}
est.lm.equal <- lm(salary.equal ~ years.equal + GM + years.equal:GM.equal)
summary(est.lm.equal)
```
  
  
### Use these questions to help you in making interpretations in this lab:

#### 1. If `x` is increased by one unit what happens to the  mean value of `y`?
a. For each year increased, a male is paid $beta3$ (which on the iteration before this was 0.1916, it changes since it is MCMC) more than a female (thousand dollars? units not known since link is broken to data).
b. E.g., in year 2, a male will make 0.1916 $\times$ 2 more than the female on average.
c. This should hover around 0. See $beta3$ normal density plot to get a sense of the potential variation. 
  
#### 2. Will the lines ever intersect over the range of the data?
a. No, the lines will never intersect, but they run nearly parallel, hence "equal" 
  
#### 3. What meaning does the slope have?
a. The slope of the interaction term means the amount more that is paid to males than females, given years of experience. 
b. The slope of years is the amount paid to males, given years of experience.
  
#### 4. You can answer Q.1 using point and interval estimates.
a. For each year increased, a male is paid 0.200 more than a female.
  
#### 5. Make sure you know how to interpret BCI's (Bayesian credible intervals) -- these are probability intervals and NOT Confidence Intervals.
a. There is a 95% probability that a male is paid 0.9383 or less than a female  (since MCMC likely to vary from above output).

## Task 3 (Equal)

  Plot the data and the estimating lines
```{r 3b, eval=TRUE}
# Plot salary on years of experience
plot(salary.equal~years.equal)

# List of coefficients from model above (classical)
cf2.equal = coef(est.lm.equal)
cf2.equal

# Plot the slope of each male and female
abline(coef = c(cf2.equal[1],cf2.equal[2]), col="darkseagreen4", lwd =3)
abline(coef = c(cf2.equal[1]+cf2.equal[3], cf2.equal[2]+cf2.equal[4]), col = "steelblue", lwd = 3)
legend("topleft",legend = c("Female","Male"),fill =c("darkseagreen4","steelblue"))
title("Male vs. Female Salaries - Equal Dataset | Daniel Carpenter")
```
  








