---
title:    "Lab 4 - Bayes Box and R"
subtitle: "Bayesian Statistics"
author:   "Daniel Carpenter"
date:     "February 2022"
fontsize: 12pt
geometry: margin=1in
output:
  # github_document:
  #   toc: yes
  #   number_sections: yes
  #   toc_depth: 2
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Task `1`: Point Estimates (*Incorrect*)

## `a.` Classical Point Estimate for 𝜃
```{r 1a, echo=TRUE, message=FALSE}
# Probability of success given N number of trials
from = 0
to   = 1
n    = 10

theta <- seq(from, to, length = n)

# Point Estimate for Theta
pointEst <- mean(theta)
pointEst
```

## `b.` Confidence Interval using 𝜃

`i.` Find classical 95% confidence interval using $\theta$
```{r 1bi, echo=TRUE, message=FALSE}
classicalEst <- pointEst + c(-1, 1) * 1.95 * sqrt(pointEst*(1 - pointEst) / n)
classicalEst
```


`ii.` Interpretation of the Interval above
```{r 1bii, echo=TRUE, message=FALSE}
print(paste0("We can be 95% confident that the true value of theta will be between ",
             round(classicalEst[1], 4), " and ", round(classicalEst[2], 4), 
             " (rounded to 4 decimal places) if repreated samples were taken."))
```

`iii.` How many sample are used to make the interval?
```{r 1biii, echo=TRUE, message=FALSE}
print(paste0("There are ", n, " number of samples used to make the interval."))
```

`iv.` How many samples are assumed to give the interpretation?
```{r 1biv, echo=TRUE, message=FALSE}
print(paste0("An infinite number of samples are assumed to give this interpretation"))
```

---

<br>

# Task `2`: Calculate/Plot Uniform Prior, Likelihood, and Post 
> Bayesian methods – Bayes’ rule. Assume discrete values of 𝜃 and a uniform prior on the values of 𝜃

## `a.` Suppose we take `theta = seq(0, 1, length = 20)`
```{r 2a, echo=TRUE, message=FALSE}
# Probability of success given N number of trials
from = 0
to   = 1
n    = 20

theta <- seq(from, to, length = n)
```

## `b.` Plot the prior using base R.
```{r colors, echo=TRUE, message=FALSE}
# Define some colors for the later plots
red   = 'tomato3'
blue  = 'steelblue'
green = 'darkseagreen'

# NOTE THAT COLORS CONSISTENT FOR PRIOR, LIK, AND POST
# BUT UNABLE TO TELL WHICH TO ASIGN BASED ON GRAPH IN LAB
colorPalette <- c(blue, green, red)
```

```{r 2b, echo=TRUE, message=FALSE}
prior = rep(1/n, n)

priorPlot <- plot(x = theta, y = prior, 
                  main = paste0("Uniform Prior on the Values of Theta | Daniel Carpenter\n n = ", n),
                  type = "b", pch = 19, col = red)
```

## `c.` Formula for the Likelihood

### Formula
$$
likelihood: p(D|\theta) = \frac{p(\theta|D)}{p(\theta) / \sum_{\theta^*}p(D|\theta^*) \ p(\theta^*)}
$$

### R Calculations
```{r 2c, echo=TRUE, message=FALSE}
# Substitute Dummy Input value for `x`
numSuccesses = round(n / 2)

likelihood  = dbinom(x=numSuccesses, size = n, prob = theta)
```

## `d.` Plot the Likelihood over Discrete Values of 𝜃
```{r 2d, echo=TRUE, message=FALSE}
likelihoodPlot <- plot(x = theta, y = likelihood, 
                       main = paste0("Likelihood over the Discrete Values of Theta | Daniel Carpenter\n",
                                     "x = ", numSuccesses, ", n = ", n),
                       type = "b", pch = 19, col = blue)
```

## `e.` Prior and the Likelihood Plotted Together
```{r 2e, echo=TRUE, message=FALSE}
# Plot the likelihood
plot(x = theta, y = likelihood, 
                     main = paste0("Likelihood and Prior over the Discrete Values of Theta | Daniel Carpenter\n",
                                   "x = ", numSuccesses, ", n = ", n),
     ylab = "Probability",
                     type = "b", pch = 19, col = blue)

# Plot the Prior
lines(x = theta, y = prior, type = "b", pch = 19, col = red)
```

## `f.` Formula and Plot of the Posterior

### Formula for Posterior
$$
posterior: p(\theta|D) = \frac{p(D|\theta) \ p(\theta)}{p(D)}
$$

### Plot of Posterior
```{r 2f, echo=TRUE, message=FALSE}
# Calculate the Prior x the Likelihood
h <-  prior * likelihood

# Get the posterior Distribution
posterior = h / sum(h)

# Plot of only the posterior
plot(x = theta, y = posterior,
     main = paste0("Posterior over the Discrete Values of Theta | Daniel Carpenter\n",
                   "x = ", numSuccesses, ", n = ", n),
     type = "b", pch = 19, col = green)
```


### Combined Plots
```{r 2f1, echo=TRUE, message=FALSE}
# Plot the Posterior
plot(x = theta, y = posterior, 
                     main = paste0("Prior, Likelihood, Posterior \nover the Discrete Values of Theta | Daniel Carpenter\n",
                                   "x = ", numSuccesses, ", n = ", n),
                     type = "b", pch = 19, col = green,
     ylab = "Probability")

# Prior
lines(x = theta, y = prior,     type = "b", pch = 19, col = red)

# Likelihood
lines(x = theta, y = likelihood, type = "b", pch = 19, col = blue)

```


---

<br>

# Task `3`: Bayes Box 

## `a.` Matrix in R that will hold the 5 column vectors
```{r 3a, echo=TRUE, message=FALSE}
# Get the length of theta
numRows <- length(theta)

bayesMatrix <- matrix(data = c(theta, prior, likelihood, h, posterior), 
                      nr = numRows, nc = 5, byrow = FALSE)
colnames(bayesMatrix) <- c("theta", "prior", "likelihood", "h", "posterior")

# Print Table
knitr::kable(bayesMatrix)

# Totals for each Column
knitr::kable(colSums(bayesMatrix))
```

## `b.` Rewrite Bayes’ theorem in terms of `h = prior * likelihood`
$$
p(\theta|D) = \frac{p(D|\theta) \ p(\theta)}{\sum_{\theta^*}p(D|\theta^*) \ p(\theta^*)}  
$$

## `c.` Plot Prior, Likelihood, and Posterior in `ggplot`
```{r 3c, echo=TRUE, message=FALSE}
## Number of Theta numbers to plot
thetaNumericValues = 1:numRows

## Convert to data fram for plotting
df <- as.data.frame(cbind(bayesMatrix, thetaNumericValues))


if(!require(tidyverse)) install.packages(tidyverse)
df <- df %>%
  
  # Select to only the needed data
  select(-h) %>%
  
  # Pivot y axis variables into single column for ggplot-ing
  pivot_longer(cols      = c("prior", "likelihood", "posterior"),
               values_to = "values",
               names_to  = "statNames")

# Plot it!
basePlot <- ggplot(df,
                   aes(x = thetaNumericValues,
                       y = values,
                       color = statNames)) +
  # Theme and colors
  theme_minimal() +
  scale_color_manual(values = colorPalette) +
  
  # Labels
  labs(title = "Prior, Likelihood, Posterior over the Discrete Values of Theta",
       subtitle = paste0("Daniel Carpenter | x = ", numSuccesses, ", n = ", n),
       x = 'Number of Theta Values', y = 'Probability') +

  # The points
  geom_point()

basePlot 

```


---

<br>

# Task `4`: Create Function `mybinpost()` to Dynamically Produce Above Calculations

## `a. - b.` Create Function `mybinpost()`
```{r 4a, echo=TRUE, message=FALSE}
mybinpost <- function(n, x, theta, alpha) {
  
  # CALCULATIONS -------------------------------------------
    
    ## Get the length of theta
    numRows <- length(theta)
  
    ## Calculate prior assuming uniform distribution
    prior = rep(1/numRows, numRows)
  
    ## Calculate the likelihood
    likelihood  = dbinom(x=x, size=n, prob=theta)
    
    ## Calculate the Prior x the Likelihood
    h <-  prior * likelihood
    
    ## Get the posterior Distribution
    posterior = h / sum(h) 
    
    ## Consolidate into a matrix with row and column names
    bayesMatrix <- matrix(c(theta, prior, likelihood, h, posterior), 
                        nr = numRows, nc = 5, byrow = FALSE)
    colnames(bayesMatrix) <- c("theta", "prior", "likelihood", "h", "posterior")
    rownames(bayesMatrix)= c(rep("", numRows))

        
  # PLOTTING -----------------------------------------------
    
    ## Number of Theta numbers to plot
    thetaNumericValues = 1:numRows
    
    ## Convert to data frame for ggplot
    df <- as.data.frame(cbind(bayesMatrix, thetaNumericValues))
    
    
    if(!require(tidyverse)) install.packages(tidyverse)
    df <- df %>%
      
      ### Select to only the needed data
      select(-h) %>%
      
      ### Pivot y axis variables into single column for ggplot-ing
      pivot_longer(cols      = c("prior", "likelihood", "posterior"),
                   values_to = "values",
                   names_to  = "statNames")
    
    ## Plot it!
    basePlot <- ggplot(df,
                       aes(x = thetaNumericValues,
                           y = values,
                           color = statNames)) +
      ### Theme and colors
      theme_minimal() +
      scale_color_manual(values = colorPalette) +
      
      ### Labels
      labs(title = "Prior, Likelihood, Posterior over the Discrete Values of Theta",
           subtitle = paste0("Daniel Carpenter | x = ", x, ", n = ", n),
           x = 'Number of Theta Values', y = 'Probability')+ 
      
      ### Create the points
      geom_point()
    
    # Output Plot
    print(basePlot)
    
    
  # BCI & Point Est ------------------------------------------
    
    ## BCI
    cp = cumsum(posterior) # cumulative sum
    L = max(which(cp<alpha/2)) # this gives the max index where  cp < alpha/2
    U = min(which(cp > 1-alpha/2))
    BCI = df$theta[c(L,U)] # close to the desired BCI
    
    ## Bayesian point estimate is the posterior mean
    bayesPointEst <- mean(posterior)
    
    
  # RETURN DATA --------------------------------------------- 
    
    ## Used for latex output of a matrix
    if(!require(xtable)) install.packages(xtable)
    
    ## Return a list of relevant data
    return(list('bayesMatrix'      = bayesMatrix,
                'latexBayesMat'    = xtable(bayesMatrix),
                'numTrials'        = n,
                'numSuccesses'     = x,
                'bayesPointEst'    = bayesPointEst,
                'bayesCredIntvl95' = BCI))
}

```


---

<br>

# Task `5`: Run `mybinpost()`

## `a.` Test #1  of `mybinpost()`
```{r 5a, echo=TRUE, message=FALSE}
mybinpost(n=20,x=12,theta = seq(0,1,length=50), alpha=0.05)
```

## `b.` Test #2 of `mybinpost()`
```{r 5b, echo=TRUE, message=FALSE}
mybinpost(n=50, x=25, theta=seq(0,1,length=30), alpha=0.1)
```

