---
title: 'Lab 11: Comparing Binomial Proportions'
author: "Wayne Stewart"
date: "April 6, 2018"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A good reference book
The following problem is taken from page 105 `Bayesian Ideas and Data Analysis: An Introduction for Scientists and Statisticians` 



# The story

534 medflies that lived at least 34 days. These are cross classified by two factors:

    A) whether the fly produced at least 1,000 eggs in the first 30 days of life and
    B) whether the fly lived ata least 44 days
    
```{r}
mat = matrix(c(54,80,224,176), nr = 2,nc=2, byrow = TRUE)
colnames(mat)= c("Long-lived","short-lived")
rownames(mat)= c("High", "Low")
mat
```
These are multinomial data but if we view the Longetivity given reproductive factor we shall see these as two independent binomials

$$y_1|\theta \sim Bin(n_1,\theta) \perp y_2|\theta_2 \sim Bin(n_2, \theta_2)$$
    
The data can then be obtained from the above table:

> 
    $n_1=134,n_2=400, y_1=54, y_2 = 224$
    
    
    
# Task 1

Find the conjugate solution -- that is find the *analytical* posteriors for $\theta_1|y_1$ and $\theta_2|y_2$ using $\theta_i\sim Beta(a_i,b_i)$ where $i\in \{1,2\}$

$$
\begin{aligned}
&p\left(\theta_{1} \mid y_{1}\right)=\operatorname{Beta}\left(a_{1}+y_{1}, b_{1}+n_{1}-y_{1}\right) \\
&p\left(\theta_{2} \mid y_{2}\right)=\operatorname{Beta}\left(a_{2}+y_{2}, b_{2}+n_{2}-y_{2}\right)
\end{aligned}
$$


# Task 2
Make plots of the two posteriors in R using

* $a_i = 1, b_i =1$
* $a_i = 4, b_i =1$
* $a_i = 1, b_i =4$
* $a_i = 3, b_i =3$

```{r postPlots, echo=TRUE, message=FALSE, warning=FALSE}
a = c(1, 4, 1, 3)
b = c(1, 1, 4, 3)
n = c(134, 400)
y = c(54, 224)

par(mfrow=c(length(n),length(a))) # layout of plots

# Plot each beta distribution
for (i in 1:length(n)) {
  for (j in 1:length(a)) {
    curve(dbeta(x, a[j]+y[i], b[j]+n[i]-y[i]),
                xlim = c(0,1), lwd=2, col = c('tomato3', 'steelblue3')[i],
                ylab = 'density')
    text(0.1, 6, paste('a=', a[j], 'b=',b[j]))
  }
}
```


# Task 3

Create a jags model for the above problem by correcting and/or completing the given code. Note that there are some useful logical nodes made. Comment on them.

```{r model, eval=FALSE}
model{
  for(i in 1:2){
  y[i] ~ dbin(theta[i], n[i])

  theta[i] ~ dbeta(1,1)             # Uniform prior now, but change to a and b params
  odds[i] <- theta[i]/(1-theta[i])  # Prob of success / prob of failure
  }
  
  RD   <- theta[2]-theta[1] # Risk difference
  RR   <-theta[2]/theta[1]  # Relative risk
  test <- step(RD)          # Probability that RD is positive
}
```



# Task 4

Now make the complete jags code. Please include some logical constant assignments so that you can assign new priors easily

Example:

```{r logicalnodes, eval=FALSE}
a[i]<-1
b[i]<-1
```


Make sure you monitor `theta`, `RR`, `OR` and `test`.
```{r rJags, echo=TRUE, message=FALSE, warning=FALSE}
require(rjags)

# alpha and beta
a = 1 
b = 1 

dataList = list(n=n, y=y, a=a, b=b) # Put variables into model's data list

modelString = "
model{
  for(i in 1:2){
  y[i] ~ dbin(theta[i], n[i])

  theta[i] ~ dbeta(a,b)             # beta distribution with hyper params a and b
  odds[i] <- theta[i]/(1-theta[i])  # Prob of success / prob of failure
  }
  
  RD   <- theta[2]-theta[1] # Risk difference
  RR   <-theta[2]/theta[1]  # Relative risk
  test <- step(RD)          # Probability that RD is positive
}
"# close quote for modelString

writeLines( modelString , con="TEMPmodel.txt" )

initsList = list(theta = c(0.4,0.5))

# Run the chains:
jagsModel = jags.model( file="TEMPmodel.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )

# Show how the model is ran - realSlicer is slow. Conjugate is fast
list.samplers(jagsModel)

update( jagsModel , n.iter=500 )
codaSamples = coda.samples( jagsModel , variable.names=c('theta', 'RR', 'RD', 'test'),
                            n.iter=33340 )
save( codaSamples , file=paste0("lab11","Mcmc.Rdata") )


# Plot the model output
require(ggmcmc)
s = ggs(codaSamples)
d = ggs_density(s); print(d)

cr = ggs_crosscorrelation(s); print(cr)
```




# Task 5

Please give point and interval estimates for the above nodes. Interpret them.
```{r sumOut, echo=TRUE, message=FALSE, warning=FALSE}
# Output of model
summary(codaSamples)
```

### Interpretation:
* Mean value of `RD` is around 0.16 with a standard deviation of around 0.049
* There is a probability of 95% that `RD` will be between around 0.06 and 0.25
* Each variable can be interpreted like I have for `RD`
* All values rounded and stated as approximately since each MCMC run will generate a slightly different estimate.

# Task 6

Make a function `mybibin()` that will solve the same type of problem where the data inputs will be a vector of y's and a vector of n's. The priors will be beta's and will therefore need a vector of a's and b's. 
The command line output should be a list of point and interval estimates for a given $\alpha$.
The function should produce analytical posterior, prior and likelihood plots and MCMC posterior plots of the theta's and all other stochastic nodes.
Run your function with the same data as given and $a = c(1,2), b= c(1,2)\; and\; \alpha = 0.2$

How will you change the model code? 
There are many ways to do this -- one way is to break the model "text" into parts and rejoin them after changing values. You can use `paste()`