---
title:  "Lab 1 - Bayesian Statistics"
author: "Daniel Carpenter"
date:   "January 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document:

    toc: yes
    toc_float: yes
    number_sections: yes
  pdf_document: 
    
    toc: TRUE
    number_sections: TRUE
  word_document:
    
    toc: TRUE
    number_secions: TRUE    
  md_document: default
     
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Tasks

## Helpful `R Markdown` Cheat sheets and Links

* [Lesson 1 for R Markdown](http://rmarkdown.rstudio.com/lesson-1.html)
* [R Markdown Cheat sheet (v2.0)](https://github.com/rstudio/cheatsheets/blob/main/rmarkdown-2.0.pdf)
* [R Markdown Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf?_ga=2.209992917.1171562397.1642623613-1554842467.1639495227)
* [Other R Markdown Cheat sheets](https://rmarkdown.rstudio.com/lesson-15.html)

---

## Examples of `LaTex` Formulae with R 

* [Note helpful LaTex repository here](https://www.overleaf.com/learn/latex/Learn_LaTeX_in_30_minutes#!#Adding_math_to_LaTeX)


<br>

#### Example `a`:
$$
(x + a)^n = 
\sum_{k=0}^{n}
  \begin{pmatrix}
    n\\
    k
  \end{pmatrix}
x^ka^{n-k}
$$


<br>

#### Example `b`:
$$
(1 + x)^n = 1 
+ \frac{nx}{1!} + \frac{n(n-1)x^2}{n!} + \dotsm
$$


<br>

#### Example `c`:
$$
f(x) = a_{0} + 
\sum_{n = 1}^{\infty}
  (a_{n}\cos\frac{n \pi x}{L} + b_{n}\sin\frac{n \pi x}{L})
$$


<br>

#### Example `d`:
$$
e^x = 1 
+ \frac{x}{1!}
+ \frac{x^2}{2!}
+ \frac{x^3}{3!}
+ \dotsm, -\infty < x < \infty
$$

---

## Histogram of $\mu$

```{r histogram, echo=TRUE, cache=TRUE}
library(rstan)
library(Intro2R)

# Suggested to do this with multiple cores
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

# Read in the ddt csv file
ddt <- read.csv("Input_Files/DDT.csv")

# Create model using rstan
basic_data <- list(y=ddt$LENGTH, N=length(ddt$LENGTH))
fit <- stan(file = "Input_Files/basic.stan",
            model_name = "basic",
            data = basic_data,
            chains = 3,
            warmup = 1000,
            cores = 3,
            iter = 5000,
            pars = c("mu")
            )

# ---

library(ggplot2)
library(rstanarm)
library(bayesplot)

# Create a blue histogram of mu
color_scheme_set("blue")
mcmc_hist(fit,pars = c("mu"))
```

---

## Assessment for Bayesian Statistics

Group             | Weight
------------------|--------
Assignments       | 40%
Midterm Exam      | 10%
Lab Exercises     | 10%
Chapter Quizzes   | 10%
Final Exam        | 30%
**Total**         | **100%**

