---
title:  "Lab 3 - Probability: Distributional Theory Continued"
subtitle: "Bayesian Statistics"
author: "Daniel Carpenter"
date:   "February 2022"
fontsize: 12pt
geometry: margin=1in
output:
  md_document: default
     
  pdf_document: 
    toc: TRUE
    number_sections: TRUE
  github_document: default
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: TRUE
    number_sections: TRUE    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Task `1`: Calculating Exact Probability

## Inputs for `mu` and `sigma`

$X \sim N(\mu = 10, \sigma = 4)$

```{r inputs, echo=TRUE, message=FALSE}
mu    = 10
sigma = 4
```

## `a.` Find 𝑃(𝑋 ≤ 8)
```{r 1a, echo=TRUE, message=FALSE}
pnorm(8, mu, sigma)
```

## `b.` Find 𝑃(𝑋 ≥ 11)
```{r 1b, echo=TRUE, message=FALSE}
1 - pnorm(11, mu, sigma)
```

## `c.` Find 𝑃(8 ≤ 𝑋 < 14)
```{r 1c, echo=TRUE, message=FALSE}
# Inputs for upper and lower bound when calculating the area
upperBound = 14
lowerBound = 8

# Calculate the area in between the two bounds
pnorm(upperBound, mu, sigma) - pnorm(lowerBound, mu, sigma)
```

## `d.` Find 𝑃(𝑋 > 10)
```{r 1d, echo=TRUE, message=FALSE}
1 - pnorm(10, mu, sigma)
```

---

<br>

# Task `2-4`: Plotting Probability

## Probability Function `Myfunction()` 
* *Please note this function is technically task 4, but is placed here for repeatability.*
```{r Myfunction, echo=TRUE, message=FALSE}
    Myfunction = function(mu, sigma, 
                          lowerBound = NA, upperBound = NA, 
                          roundTo = 4, 
                          color = paste0("lightsteelblue", floor(runif(1, min=1, max=4))),
                          returnCMD = FALSE)
    {
      
      # Calculate the "xlim" lower and upper bound for the Normal PDF Curve
      curveLowerBound <- mu - 3*sigma
      curveUpperBound <- mu + 3*sigma
      
      # Initialize variables related to output and graph
      title <- ""   # Title of graph
      exactProb = 0 # The exact probability of the questions
      
      # If no provided LOWER AND UPPER Bound (NA as parameter value) then assume none
      if (!(is.na(lowerBound)) & !(is.na(upperBound))) {
        title <- paste0(", P(", lowerBound, " <= X < ",upperBound,")")
        exactProb = pnorm(upperBound, mu, sigma) - pnorm(lowerBound, mu, sigma) # calculate prob 
        
      # If no provided LOWER Bound (NA as parameter value) then assume none
      } else if (is.na(lowerBound)) {
        lowerBound = curveLowerBound
        title <- paste0(", P(X < ",upperBound,")") # Set a dynamic title
        exactProb = pnorm(upperBound, mu, sigma) # calculate prob 
        
      # If no provided UPPER Bound (NA as parameter value) then assume none
      } else if(is.na(upperBound)) {
        upperBound = curveUpperBound
        title <- paste0(", P(X >= ",lowerBound,")")
        exactProb = 1 - pnorm(lowerBound, mu, sigma) # calculate prob 
      }
      
      # Create the line that displays the bell curve (between the CURVE bounds defined above)
      curve(
        
        ## Normally Distributed
        dnorm(x,mu,sigma), 
        
        ## Normally Distributed
        xlim=c(curveLowerBound, curveUpperBound), 
        
        ## Line width
        lwd =2, 
        
        ## Title with descriptive characteristics about function parameters
        main = paste0("Probability Distribution (by Daniel Carpenter)\n",
                      "X ~ N(",mu,", ",sigma,")", title)
      )
      
      # Add the AREA of between the lower and upper bound P(lowerBound<X<=upperBound)
      
        ## X-Axis curve (length does not matter)
        xcurve = seq(lowerBound,upperBound, length=1000)
        
        ## Y-Axis Curve
        ycurve = dnorm(xcurve, mu,sigma)
        
        ## Combine the X and Y curve to form the area (in green)
        polygon(c(lowerBound, xcurve, upperBound), 
                c(0, ycurve, 0), 
                col=color) 
      
      # Add the probability as text
        
        ## Calculate the area (probability)
        area = exactProb
        areaRounded = round(area, roundTo)
        
        ## Place this on the above plot
        text(12,0.02,substitute(paste("Probability = ", areaRounded), 
                                list(areaRounded = areaRounded)))
      
      # Return stats about the Plot
      if (returnCMD) {
        return(list(mu = mu, 
                    sigma = sigma,
                    prob = areaRounded))
      }
    }
```

## `a.` Find 𝑃(𝑋 ≤ 8)
```{r 2a, echo=TRUE, message=FALSE}
Myfunction(mu, sigma, upperBound = 8)
```

## `b.` Find 𝑃(𝑋 ≥ 11)
```{r 2b, echo=TRUE, message=FALSE}
Myfunction(mu, sigma, lowerBound = 11)
```

## `c.` Find 𝑃(8 ≤ 𝑋 < 14)
```{r 2c, echo=TRUE, message=FALSE}
Myfunction(mu, sigma, lowerBound = 8, upperBound = 14)
```

## `d.` Find 𝑃(𝑋 > 10)
```{r 2d, echo=TRUE, message=FALSE}
Myfunction(mu, sigma, lowerBound = 10)
```

---

<br>

# Task `5`: Create and Demonstrate Function `mynormplot()`

## `a-f` Create `mynormplot()` Function
```{r mynormplot, echo=TRUE, message=FALSE}
# Get function from above since it is dynamic and can handle the below calculations
mynormplot <- Myfunction
```

## `g.` Calculate 𝑃(7 ≤ 𝑋 ≤ 10), where 𝑋 ∼ 𝑁(8,5) with `mynormplot()` Function
```{r 5g, echo=TRUE, message=FALSE}
mynormplot(
  
  # Stats for Normal Distribution Creation
  mu = 8, 
  sigma = 5, 
  
  # Bounds for Calculating Probability
  lowerBound = 7, 
  upperBound = 10, 
  
  # Round to 6 decimal places
  roundTo = 6, 
  
  # Set color of the prob. area to "Blue"
  color = "Blue",
  
  # Since TRUE, we will return command line output with
  # mean of the normal, the standard deviation and the probability calculated
  returnCMD = TRUE) 
```

---

<br>

# Task `6`: 
Formula for Normal Density using $\LaTeX: \ \ X \sim N(\mu, \sigma)$

---

<br>

# Task `7`:
```{r ggplot, echo=TRUE, message=FALSE}
if(!require(ggplot2)) install.packages("ggplot2")

ggplot(data = NULL, aes(c(-10,10))) +
  
  # LHS of the distribution (Red)
  geom_area(stat = "function", 
            fun = dnorm, 
            args = list(mean = 0, sd = 3),
            fill = "red", xlim = c(-10, 0))  + 
  
  # RHS of the distribution (Blue)
  geom_area(stat = "function", 
            fun = dnorm, 
            args = list(mean = 0, sd = 3),
            fill = "blue", xlim = c(0, 10))  + 
  
  # Labels on the chart
  labs(title = "Daniel Carpenter",
       subtitle = "Normal Distribution with ggplot")
```
