---
title:    "Assignment 3"
subtitle: "Bayesian Statistics"
author:   "Daniel Carpenter"
date:     "April 2022"
fontsize: 12pt
geometry: margin=1in
output:
  html_document:
    toc: yes
    toc_float: yes
  # github_document:
  #   toc: yes
  #   # number_sections: yes
  #   toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

---

# Task 1

## `a`-`b`

```{r}
# The data input
y = c( rep(1,9),rep(0,3) , rep(1,45),rep(0,15) , rep(1,3),rep(0,9) ) 
s = c( rep("A",12) , rep("B",60) , rep("C",12) )
fileName = "Ass3.1.csv"
write.csv( data.frame(y=y,s=s) , file=fileName , row.names=FALSE )


# Example for Jags-Ydich-XnomSsubj-Mbernbeta.R 
#------------------------------------------------------------------------------- 
# Optional generic preliminaries:
graphics.off() # This closes all of R's graphics windows.

#------------------------------------------------------------------------------- 
# Load The data 
myData = read.csv(fileName)

# Include head of data
head(myData)

# N.B.: The functions below expect the data to be a data frame, 
# with one component named y being a vector of integer 0,1 values,
# and one component named s being a factor of subject identifiers.
myData$s = factor( myData$s )

#------------------------------------------------------------------------------- 
# Load the relevant model into R's working memory:
source("Jags-Ydich-XnomSsubj-MbernBeta.R")

#------------------------------------------------------------------------------- 
# Optional: Specify filename root and graphical format for saving output.
# Otherwise specify as NULL or leave saveName and saveType arguments 
# out of function calls.
dir.create('Output')
fileNameRoot = "Output//"
graphFileType = "png" 

#------------------------------------------------------------------------------- 
# Generate the MCMC chain:
mcmcCoda = genMCMC( data=myData , numSavedSteps=50000 , saveName=fileNameRoot )

#------------------------------------------------------------------------------- 
# Display diagnostics of chain, for specified parameters:
parameterNames = varnames(mcmcCoda) # get all parameter names
for ( parName in parameterNames ) {
  diagMCMC( codaObject=mcmcCoda , parName=parName , 
            saveName=fileNameRoot , saveType=graphFileType )
}

# Ensure that the graphs of theta[1] and [2] are shown
knitr::include_graphics(paste0(fileNameRoot, 'Diagtheta[1].', graphFileType), dpi = 5)
knitr::include_graphics(paste0(fileNameRoot, 'Diagtheta[2].', graphFileType), dpi = 5)

#------------------------------------------------------------------------------- 
# Get summary statistics of chain:
summaryInfo = smryMCMC( mcmcCoda , compVal=NULL , rope=c(0.45,0.55) ,
                        compValDiff=0.0 , ropeDiff = c(-0.05,0.05) ,
                        saveName=fileNameRoot )

# Display posterior information:
plotMCMC( mcmcCoda , data=myData , compVal=NULL , rope=c(0.45,0.55) ,
          compValDiff=0.0 , ropeDiff = c(-0.05,0.05) ,
          saveName=fileNameRoot , saveType=graphFileType )
```


## `c` Credibility of Model
* The MCMC estimation appears credible across all values of theta since:
  * Param. Values seem to be around the mean value, which is stationary
  * Shrink factor evens out
  * Autocorrelation drops quickly
  * Each chain's desnity represents each other well when viewing superimposed over each other

## `d` Sample Size Effect
* Increased sample size of subject B will cause for a tighter distribution on the estimation

```{r}
# Add the histogram
plotPost(mcmcCoda[,'theta[1]'], xlim = c(0,1), xlab= expression(theta[1]),
         main = 'Daniel Carpenter', col = 'darkseagreen3',
                      showCurve = FALSE)

# Add the density line
paramSampleVec = as.matrix(mcmcCoda[,'theta[1]'])
densCurve = density( paramSampleVec , adjust=2 )
a = densCurve$x; b = densCurve$y

lines(a, b, type="l" , lwd=2.5 , col='darkseagreen4' , bty="n")


graphics.off() # This closes all of R's graphics windows.
```

## `e`-`f` Lines()
* `a` and `b` resemble come from the `density()` function from the `mcmcCoda[,'theta[1]']` object
* The lines command adds a line to a plot as a layer.

## `h` Functions
i. `xlab= expression(theta[1])`
ii. `col = 'green3'`; note I used `col = 'darkseagreen3'`
iii. `main = 'Daniel Carpenter'`

## `i` Lines()
* You can use `showCurve = TRUE` but it does not plot both the histogram and the line, so I manually added the line.

